---
title: "Reproducible Research: Peer Assessment 1"
author: "Carlor Roberto"
date: "Wednesday, November 12, 2014"
output: html_document
keep_md: true
---

This project makes use of data from a personal activity monitoring device.
This device collects data at 5 minute intervals through out the day. The data
consists of two months of data from an anonymous individual collected during
the months of October and November, 2012 and include the number of steps
taken in 5 minute intervals each day

### load libraries
```{r}
library(lubridate)
library(plyr)
library(dplyr)
library(sqldf)
```

## Loading and preprocessing the data
```{r echo=TRUE}

df <- read.csv(file = "activity.csv", header = TRUE, na.strings = "NA", sep = ",")

str(df)
```
### filter invalid data
```{r echo=TRUE}
dfna <- na.omit(df)
str(dfna)
```
### plot histogram of the total number of steps taken each day
```{r echo=TRUE}
dfstep <- dfna[,c("steps","date","interval")]
dfstep$date <- ymd(dfstep$date)
with(dfstep, plot(steps ~ date, type="h"))
```

## What is mean total number of steps taken per day?
Mean and median can be calculated using the mean() and median() functions

```{r echo=TRUE}
df_mean <- ddply(dfna, ~ date, summarise, mean = mean(steps), median = median(steps))
print(df_mean)
```

## What is the average daily activity pattern?
Once we have the vaerage per day, it is easy to calculate the average daily
```{r echo=TRUE}
mean(df_mean$mean)
```
### plot the steps against the 5-minute interval
```{r echo=TRUE}
histdata <- with(dfstep, plot(steps ~ interval, type="h"))
```
Checking the graphyc we can see that the 5-minute interval with maximum number of steps is 7:20 am

## Imputing missing values
```{r echo=TRUE}
nas <- is.na(df[,1])
count.na <- table(nas)[2]
```
1. Number of missing NAs
```{r echo=TRUE}
print(count.na)
```
2. A good approach to fill NAs is to use the mean calculated in df_men

3. create a new dataset based on fill NAs
```{r echo=TRUE}
newdf <- df
getdate=NULL
getmean=df[1,]
for(i in 1:nrow(df)){
    if(is.na(df[i,1]) && (df[i,"date"] == "2012-10-01" || df[i,"date"] == "2012-10-08" || df[i,"date"] == "2012-11-01" || df[i,"date"] == "2012-11-04" || df[i,"date"] == "2012-11-09" || df[i,"date"] == "2012-11-10" || df[i,"date"] == "2012-11-14" || df[i,"date"] == "2012-11-30" )){
       newdf[i,1] <- 0
       next
    }    
    # now fill the remaining with mean value    
    if(is.na(df[i,1])){
        getdate <- as.character(df[i,"date"])
        m=0
        for(j in 1:nrow(df_mean)){
            if(as.character(df_mean[j,"date"]) == getdate){
                m = df_mean[j,"mean"]    
                break
            }
        }
        # fill with mean
        #newdf[i,"steps"] <- m    
        
    }
}
```
4. No, as we can see in the histogram, there is no impact of imputting missing as we replaced NA by zero values in dataframe.

```{r echo=TRUE}
ndf_mean <- ddply(newdf, ~ date, summarise, mean = mean(steps), median = median(steps))
#print(ndf_mean)
mean(ndf_mean$mean)
#nhistdata <- with(dfstep, plot(steps ~ interval, type="h"))
ndfstep <- newdf[,c("steps","date","interval")]
ndfstep$date <- ymd(ndf_mean$date)
with(ndfstep, plot(steps ~ interval, type="h"))
```

## Are there differences in activity patterns between weekdays and weekends?
Yes, we can see that there are more activity during weekdays than weekends

```{r}
w.ndfstep <- ndfstep
w.ndfstep["week"] <- weekdays(as.Date(ndfstep$date), abbreviate=T)
for(i in 1:nrow(w.ndfstep)){
    if(w.ndfstep[i,"week"] == "sab" || w.ndfstep[i,"week"] == "dom" )
        w.ndfstep[i,"week"] = "weekend"
    else
        w.ndfstep[i,"week"] = "weekday"
}
w.ndfstep$wee = as.factor(w.ndfstep$wee)
```
# load the library and plot
```{r}
library(lattice)
xyplot(steps ~ interval | week, data = w.ndfstep, layout = c(1,2), type = "l")

```



